---
title: "Reweighting with `reweight`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reweighting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(tibble.print_min = 4L, tibble.print_max = 4L)
```

```{r setup, include = FALSE}
library(microweight)
library(tidyverse)
library(knitr)
```


"Reweighting" is the process of adjusting the weights on a microdata file so that the file, with new weights, better represents the population of interest.

The function `reweight` creates these new weights.

## Setting up our data

In this vignette we'll use small microdata files extracted from the American Community Survey in the United States: `acs` (10,000 individuals) and `acsbig` (100,000 individuals).

To keep things simple, we create a minimal subset with only these columns:

Column        | Description
------------- | -------------
pwgtp         | person weight
agep          | person's age
mar           | marital status (1=married, 2=single, ...)
pincp         | personal income, total
wagp          | wage income
ssp           | Social Security income

The first few records have a lot of zeros so we'll look at the tail, which is more representative of typical records:

```{r}
acs2 <- acs %>%
  select(pwgtp, agep, mar, pincp, wagp, ssp)
tail(acs2)
```

Look at weighted totals for a few variables:
```{r}
totals <- acs2 %>%
  summarise(across(c(pincp, wagp, ssp), ~sum(. * pwgtp)))
totals
```

## A simple reweighting problem

Suppose that we have other information, perhaps for a later year, and that we want the following targets:

Column        |   Current value  | Targeted value
------------- | -----------------|----------------
pincp         |    7,990,814,687 |    8,000,000,000
wagp          |    6,070,651,037 |    6,020,123,456
ssp           |      546,119,219 |      550,100,000

We want new weights that will hit these targets. To do that we must define the following:

Item          |   Description
------------- | -------------
iweights      | vector of initial weights, here pwgtp
target_names  | vector of names of the target variables
targets       | vector of values for the targets
tol           | vector of tolerances around the target values
xmat          | matrix with values

### Set up the problem

Let's define the first four, as they are easy to understand:
```{r}
iweights <- acs2$pwgtp
target_names <- c("pincp", "wagp", "ssp")
targets <- c(8e9, 6020123456, 550.1e6)
tol <- c(100, 100, 100)
```


`xmat` is simply a matrix of values:
```{r}
xmat <- acs2 %>%
  select(all_of(target_names)) %>%
  as.matrix()
```


### Solve the problem
Now we are ready to reweight:
```{r}
result <- reweight(iweights, targets, target_names, tol, xmat)
```

### Examine results
How did it work and how long did it take (in seconds)?

```{r}
result$solver_message
result$etime
```

How do the targeted values compare to the original values? `reweight` returns a list with target_df that includes the following information:


Item            |   Description
--------------- | -------------
targnum         | target number
targname        | target name
target          | targeted value
targinit	      | value for the target using initial weights
targcalc        | value calculated with new weights
targinit_diff	  | difference between target and initial value
targtol	        | tolerance allowed around the target
targcalc_diff	  | difference between calculated value with new weights and target
targinit_pdiff	| percent difference between target and initial value
targtol_pdiff	  | percent difference allowed, based upon tolerances
targcalc_pdiff  | percent difference between calculated value and target


```{r}
result$targets_df %>%
  kable(digits=c(rep(0, 8), 2, 2, 2),
        format.args=list(big.mark = ","))
```

Note that the actual difference between the target for `wagp` (wages) was $104 even though we set a tolerance of $100. That is because the optimizer that chooses new weights (in this case) has its own set of tolerances. Those can be changed by passing options to the solver, but we won't do that in this vignette.

